---
title: "Charges EDA"
author: "Connor King"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(ggExtra))
suppressMessages(library(tidyverse))
suppressMessages(library(ggpubr))
```

```{r}
support2 <- read_csv("C:/Users/Conno/Documents/Career/Projects/Hospital_Charges/data/support2.csv")
```
# Distribution Plots

```{r}
# png("./eda_plots/box_charges.png", width = 1000, height = 600)

boxplot(support2$charges,
        main = "Hospital Charges Box Plot",
        xlab = "Hospital Charges",
        horizontal = TRUE,
        col = "skyblue")

# dev.off()

```


```{r}
ggplot(data = support2, aes(x = charges)) +
  geom_boxplot(fill = "skyblue") +
  labs(
    title = "Hospital Charges Box plot",
    x = "Hospital Charges"
  ) +
  theme_bw()

# ggsave(filename = "box_charges.png", path = "./eda_plots", bg = "white")
```



```{r}
support2 %>% 
  filter(charges < 250000) %>% 
  ggplot(aes(x = charges, y = factor(dzgroup), fill = factor(dzclass))) +
    geom_violin() +
    labs(
      title = "Hospital Charges Violin plot (< $250,000)",
      x = "Hospital Charges",
      y = "Dissease Sub Category",
      fill = "Disease Category"
    ) +
    theme_bw()

ggsave(filename = "violin_charges_diseases.png", path = "./eda_plots", bg = "white")
```


Clearly, there are a significant amount of outliers.

```{r}
ggplot(data = support2, aes(x = charges, y = factor(dzgroup), fill = factor(dzclass))) +
  geom_boxplot() +
  labs(
    title = "Hospital Charges Box plot",
    x = "Hospital Charges",
    y = "Dissease Sub Category",
    fill = "Disease Category"
  ) +
  theme_bw()

ggsave(filename = "box_charges_diseases.png", path = "./eda_plots", bg = "white")
```
```{r}
ggplot(data = support2, aes(x = slos, y = factor(dzgroup), fill = factor(dzclass))) +
  geom_boxplot() +
  labs(
    title = "Hospital Charges Box plot",
    x = "Days in Study",
    y = "Dissease Sub Category",
    fill = "Disease Category"
  ) +
  theme_bw()

# ggsave(filename = "box_charges_diseases.png", path = "./eda_plots", bg = "white")
```



```{r}
support2 %>% 
  ggplot(aes(x = charges)) +
  geom_histogram(fill = "blue", color = "black") +
  labs(x = "Hospital Charges", y = "Count", title = "Hospital Charges Histogram") +
  theme_bw()
ggsave(filename = "histogram_charges.png", path = "./eda_plots", bg = "white")
```

# Scatter Plots

```{r}
charges_slos_scatter <- support2 %>% 
        ggplot(aes(x = slos, y = charges)) + 
        geom_point(color = "blue", alpha = 0.2) +
        geom_smooth(method = lm, color = "red") +
        stat_cor(p.accuracy = 0.001, r.accuracy = 0.01) +
        labs(
          y = "Hospital Charges",
          x = "Days from Study Entry to Discharge",
          title = "Days in Study and Hospital Charges"
        ) +
        theme_bw() 

f_charges_slos_scatter <- ggMarginal(
  charges_slos_scatter,
  type = "histogram",
  fill = "blue",
  color = "black"
)

f_charges_slos_scatter
ggsave(filename = "scatter_charges_slos.png", path = "./eda_plots", bg = "white")
```


```{r}
charges_hday_scatter <- support2 %>% 
        ggplot(aes(x = hday, y = charges)) + 
        geom_point(color = "blue", alpha = 0.2) +
        geom_smooth(method = lm, color = "red") +
        stat_cor(p.accuracy = 0.001, r.accuracy = 0.01) +
        labs(
          y = "Hospital Charges",
          x = "Days in Hospital at study entry",
          title = "Days in Hospital at study entry and Hospital Charges"
        ) +
        theme_bw()

ggMarginal(
  charges_hday_scatter,
  type = "histogram",
  fill = "blue",
  color = "black"
)

ggsave(filename = "scatter_charges_hday.png", path = "./eda_plots", bg = "white")
```

```{r}
charges_avtisst_scatter <- support2 %>% 
        ggplot(aes(x = avtisst, y = charges)) + 
        geom_point(color = "blue", alpha = 0.2) +
        geom_smooth(method = lm, color = "red") +
        stat_cor(p.accuracy = 0.001, r.accuracy = 0.01) +
        labs(
          y = "Hospital Charges",
          x = "TISS Score",
          title = "Days of follow-up and Hospital Charges"
        ) +
        theme_bw()

f_charges_actisst_scatter <- ggMarginal(
  charges_avtisst_scatter,
  type = "histogram",
  fill = "blue",
  color = "black"
)

f_charges_actisst_scatter

ggsave(filename = "scatter_charges_avtisst.png", path = "./eda_plots", bg = "white")
```


# Quantitative EDA

```{r}
summary(support2$charges)
```


```{r}
support2 %>%
  group_by(dzclass) %>% 
  summarize(mean_charges = mean(charges, na.rm = TRUE), mean_slos = mean(slos, na.rm = TRUE))
```

```{r}
support2 %>% 
  filter(charges > 1000000) %>% 
  select(charges, slos, hday, everything()) %>%
  arrange(desc(charges))
```

The outliers don't seem to scream anything out of the ordinary in regards to the other variables so mis-input is unlikely.


```{r}
numeric_cols <- support2 %>% select_if(is.numeric)

corr_charges <- sapply(numeric_cols, function(x) cor(support2$charges, x, use = 'complete.obs'))

sort(corr_charges)
```

Ignoring the `totmcst` and `totcst` columns, there is no discernible pattern behind the egregious outliers. Thus, we will trim the `support2` dataset 
